---
title: "heathland_fire_alpha_diversity"
author: "pgfarsund"
date: "`r Sys.Date()`"
output: html_document
---

this script performs calculations, statistical testing, and plotting for alpha diversity analysis included in Rittl et al. 2024. every chunk of code includes one organismal group and performs the following tasks: 
1) calculate richness and shannon diversity in all samples
2) test if richness or shannon diversity is different between the three successional phases
3) make boxplots showing alpha diversity across phases

```{r}
library(here)
library(tidyverse)
library(phyloseq)
library(ggpubr)
library(janitor)
library(emmeans)
library(broom)

set.seed(123)
```


plants

```{r}
# load phyloseq object:
plants <- readRDS(here("data/clean_data/heathland-fire-plants.RDS"))

# calculate alpha diversity measures and find medians and IQR:
(tibble(
  estimate_richness(plants,
    measures = c("Observed", "Shannon")
  ),
  data.frame(plants@sam_data)
) |>
  select(Observed, Shannon, phase) |>
  group_by(phase) |>
  mutate(
    group = "Plants",
    N = vctrs::vec_count(phase)$count,
    q1_richness = quantile(Observed)[2],
    med_richness = quantile(Observed)[3],
    q3_richness = quantile(Observed)[4],
    q1_shannon = quantile(Shannon)[2],
    med_shannon = quantile(Shannon)[3],
    q3_shannon = quantile(Shannon)[4],
    richness = paste0(med_richness, " (IQR ", round(q3_richness-q1_richness, digits = 2), ")"), 
    shannon = paste0(round(med_shannon, digits = 2), " (IQR ", round(q3_shannon-q1_shannon, digits = 2), ")")
  ) |>
  select(group, phase, richness, shannon, N) |>
  distinct() |>
  arrange(phase = factor(
    phase,
    levels = c("P", "B", "M")
  )) -> alpha_stats_plants)

# test whether differences in alpha diversity are significant between phases:
tibble(
  estimate_richness(plants,
    measures = c("Observed", "Shannon")
  ),
  data.frame(plants@sam_data)
) -> df

# richness

# we used GLM to test if richness was significantly different between successional phases. because richness is count data, a poisson distribution is appropriate. however, if the variance is larger than the mean richness, the data are overdispersed and we should use quasipoisson
if (var(df$Observed) > mean(df$Observed)) {
  print("Data is overdispersed, use quasipoisson distribution")
} else {
  print("Data is not overdispersed, use poisson distribution")
}

# run the GLM
(plants_rich_glm <- data.frame(tidy(glm(Observed ~ phase, data = df, family = poisson))) |> 
  mutate(group = "Plants", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3)))

# we got a significant result, therefore we perform a pairwwise comparison test with emmeans
(plants_rich_pairwise <- data.frame(emmeans(glm(Observed ~ phase, data = df, family = poisson), 
                                specs = pairwise ~ phase)$contrast) |> 
  mutate(group = "Plants", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3)))

# shannon diversity

# we test differences in Shannon diversity with a GLM. Shannon diversity is a continuous variable, so we basically want to use run the GLM with a gaussian distribution. we first check that the model's errors are not significantly different from a normal distribution using Shapiro-Wilk test

if (shapiro.test(rstandard(glm(Shannon ~ phase, data = df, family = gaussian)))$p.value >= 0.05) {
  print("The model's errors are not significanty different from a normal distribution, continue with gaussian distribution")
} else {
  print("The model's errors are significanty different from a normal distribution, decide on another distribution")
}

(plants_sha_glm <- data.frame(tidy(glm(Shannon ~ phase, data = df, family = gaussian))) |> 
  mutate(group = "Plants", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3))) 

# we got a significant result, therefore we perform a pairwwise comparison test with emmeans
(plants_sha_pairwise <- data.frame(emmeans(glm(Shannon ~ phase, data = df, family = gaussian), 
                                specs = pairwise ~ phase)$contrast) |> 
  mutate(group = "Plants", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3)))

df |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Observed,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(color = "white"),
    axis.ticks.x = element_blank()
  ) +
  scale_fill_brewer(palette = "Greens") +
  scale_color_manual(values = rep("black", times = length(unique(plants@sam_data$phase)))) +
  ylab("Richness") +
  ggtitle("Plants") +
  geom_segment(
    x = 1, xend = 2,
    y = 25, yend = 25
  ) +
  annotate(
    geom = "text", label = "**", size = 6,
    x = 1.5, y = 25.5
  ) +
  geom_segment(
    x = 1, xend = 3,
    y = 23, yend = 23
  ) +
  annotate(
    geom = "text", label = "*", size = 6,
    x = 2, y = 23.5
  ) +
  coord_cartesian(ylim = c(9, 27)) -> plarich
plarich

df |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Shannon,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_fill_brewer(palette = "Greens") +
  scale_color_manual(values = rep("black", times = length(unique(plants@sam_data$phase)))) +
  ylab("Shannon") +
  ggtitle(" ") +
  geom_segment(
    x = 1, xend = 2,
    y = 3.25, yend = 3.25
  ) +
  annotate(
    geom = "text", label = "**", size = 6,
    x = 1.5, y = 3.3
  ) +
  geom_segment(
    x = 1, xend = 3,
    y = 3, yend = 3
  ) +
  annotate(
    geom = "text", label = ".", size = 6,
    x = 2, y = 3.2
  ) +
  coord_cartesian(ylim = c(1.5, 3.4)) -> plasha
plasha
```


malaise traps

```{r}
# load and normalise phyloseq object:
malaise <- readRDS(here("data/clean_data/heathland-fire-malaise.RDS"))
malaise <- prune_taxa(taxa_sums(malaise) > 0, malaise)

malrar <- rarefy_even_depth(malaise,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = min(sample_sums(malaise))
)

# calculate alpha diversity measures and find medians and IQR:
(tibble(
  estimate_richness(malrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(malrar@sam_data)
) |>
  select(Observed, Shannon, phase) |>
  group_by(phase) |>
  mutate(
    group = "Malaise",
    N = vctrs::vec_count(phase)$count,
    q1_richness = quantile(Observed)[2],
    med_richness = quantile(Observed)[3],
    q3_richness = quantile(Observed)[4],
    q1_shannon = quantile(Shannon)[2],
    med_shannon = quantile(Shannon)[3],
    q3_shannon = quantile(Shannon)[4],
    richness = paste0(med_richness, " (IQR ", round(q3_richness-q1_richness, digits = 2), ")"), 
    shannon = paste0(round(med_shannon, digits = 2), " (IQR ", round(q3_shannon-q1_shannon, digits = 2), ")")
  ) |>
  select(group, phase, richness, shannon, N) |>
  distinct() |>
  arrange(phase = factor(
    phase,
    levels = c("P", "B", "M")
  )) -> alpha_stats_malaise)

# test whether differences in alpha diversity are significant between phases:
tibble(
  estimate_richness(inrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(inrar@sam_data)
) -> df

# richness
mod <- glm(Observed ~ phase, data = df)
summary(mod)
emmeans(mod, pairwise ~ phase)

# shannon
hist(df$Shannon, breaks = 30)
shapiro.test(df$Shannon)
mod <- glm(Shannon ~ phase, data = df)
summary(mod)
emmeans(mod, pairwise ~ phase)

df |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Observed,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_point(size = 2, shape = 21, stroke = 1, alpha = 0.8) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(color = "white"),
    axis.ticks.x = element_blank()
  ) +
  scale_fill_manual(values = c("#FDE0EF", "#DE77AE", "#C51B7D")) +
  scale_color_manual(values = rep("black", times = length(unique(inrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle("Malaise traps") -> inrich
inrich

df |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Shannon,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_point(size = 2, shape = 21, stroke = 1, alpha = 0.8) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    plot.title = element_text(color = "white")
  ) +
  scale_fill_manual(values = c("#FDE0EF", "#DE77AE", "#C51B7D")) +
  scale_color_manual(values = rep("black", times = length(unique(inrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle(" e") -> insha
insha

ggarrange(inrich, insha, nrow = 2) -> malaise_alphaplot
malaise_alphaplot
```


pitfall traps

```{r}
# load and normalise phyloseq object:
pit <- readRDS(here("data/clean_data/heathland-fire-pitfall.RDS"))

pitrar <- rarefy_even_depth(pit,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = 10
)

# calculate alpha diversity measures and find medians and IQR:
(tibble(
  estimate_richness(pitrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(pitrar@sam_data)
) |>
  select(Observed, Shannon, phase) |>
  group_by(phase) |>
  mutate(
    group = "Pitfall",
    N = vctrs::vec_count(phase)$count,
    q1_richness = quantile(Observed)[2],
    med_richness = quantile(Observed)[3],
    q3_richness = quantile(Observed)[4],
    q1_shannon = quantile(Shannon)[2],
    med_shannon = quantile(Shannon)[3],
    q3_shannon = quantile(Shannon)[4],
    richness = paste0(med_richness, " (IQR ", round(q3_richness-q1_richness, digits = 2), ")"), 
    shannon = paste0(round(med_shannon, digits = 2), " (IQR ", round(q3_shannon-q1_shannon, digits = 2), ")")
  ) |>
  select(group, phase, richness, shannon, N) |>
  distinct() |>
  arrange(phase = factor(
    phase,
    levels = c("P", "B", "M")
  )) -> alpha_stats_pitfall)

# test whether differences in alpha diversity are significant between phases:
tibble(
  estimate_richness(groundrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(groundrar@sam_data)
) -> df

# richness

# we used GLM to test if richness was significantly different between successional phases. because richness is count data, a poisson distribution is appropriate. however, if the variance is larger than the mean richness, the data are overdispersed and we should use quasipoisson
if (var(df$Observed) > mean(df$Observed)) {
  print("Data is overdispersed, use quasipoisson distribution")
} else {
  print("Data is not overdispersed, use poisson distribution")
}

# run the GLM
(pit_rich_glm <- data.frame(tidy(glm(Observed ~ phase, data = df, family = poisson))) |>
  mutate(group = "Pitfall", .before = 1)) |> 
  mutate(p.value = round(p.value, digits = 3)) 

# we did not got a significant result, therefore we did not perform a pairwwise comparison test

# shannon diversity

# we test differences in Shannon diversity with a GLM. Shannon diversity is a continuous variable, so we basically want to use run the GLM with a gaussian distribution. we first check that the model's errors are not significantly different from a normal distribution using Shapiro-Wilk test

if (shapiro.test(rstandard(glm(Shannon ~ phase, data = df, family = gaussian)))$p.value >= 0.05) {
  print("The model's errors are NOT significanty different from a normal distribution, continue with gaussian distribution")
} else {
  print("The model's errors ARE significanty different from a normal distribution, decide on another distribution")
}

(pit_sha_glm <- data.frame(tidy(glm(Shannon ~ phase, data = df, family = gaussian))) |>
  mutate(group = "Pitfall", .before = 1)) |> 
  mutate(p.value = round(p.value, digits = 3)) 

# we did not get a significant result, therefore we did not perform a pairwise comparison test

df |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Observed,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(color = "white"),
    axis.ticks.x = element_blank()
  ) +
  scale_fill_manual(values = c("tan", "tan3", "tan4")) +
  scale_color_manual(values = rep("black", times = length(unique(groundrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle("Pitfall traps") -> grorich
grorich

df |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Shannon,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_fill_manual(values = c("tan", "tan3", "tan4")) +
  scale_color_manual(values = rep("black", times = length(unique(groundrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle(" ") -> grosha
grosha

```

nematodes
```{r}
# load and normalise phyloseq object:
nematodes <- readRDS(here("data/clean_data/heathland-fire-nematodes.RDS"))

nemrar <- rarefy_even_depth(nematodes,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = min(sample_sums(nematodes))
)

# calculate alpha diversity measures and find medians and IQR:
(tibble(
  estimate_richness(nemrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(nemrar@sam_data)
) |>
  select(Observed, Shannon, phase) |>
  group_by(phase) |>
  mutate(
    group = "Nematodes",
    N = vctrs::vec_count(phase)$count,
    q1_richness = quantile(Observed)[2],
    med_richness = quantile(Observed)[3],
    q3_richness = quantile(Observed)[4],
    q1_shannon = quantile(Shannon)[2],
    med_shannon = quantile(Shannon)[3],
    q3_shannon = quantile(Shannon)[4],
    richness = paste0(med_richness, " (IQR ", round(q3_richness-q1_richness, digits = 2), ")"), 
    shannon = paste0(round(med_shannon, digits = 2), " (IQR ", round(q3_shannon-q1_shannon, digits = 2), ")")
  ) |>
  select(group, phase, richness, shannon, N) |>
  distinct() |>
  arrange(phase = factor(
    phase,
    levels = c("P", "B", "M")
  )) -> alpha_stats_nematodes)

# test whether differences in alpha diversity are significant between phases:
tibble(
  estimate_richness(nemrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(nemrar@sam_data)
) -> df

# richness

# we used GLM to test if richness was significantly different between successional phases. because richness is count data, a poisson distribution is appropriate. however, if the variance is larger than the mean richness, the data are overdispersed and we should use quasipoisson
if (var(df$Observed) > mean(df$Observed)) {
  print("Data is overdispersed, use quasipoisson distribution")
} else {
  print("Data is not overdispersed, use poisson distribution")
}

# run the GLM
(nema_rich_glm <- data.frame(tidy(glm(Observed ~ phase, data = df, family = poisson))) |>
  mutate(p.value = round(p.value, digits = 3)) |> 
  mutate(group = "Nematodes", .before = 1))

# we did not get a significant result, therefore we did not perform a pairwwise comparison test 

# shannon diversity

# we test differences in Shannon diversity with a GLM. Shannon diversity is a continuous variable, so we basically want to use run the GLM with a gaussian distribution. we first check that the model's errors are not significantly different from a normal distribution using Shapiro-Wilk test

if (shapiro.test(rstandard(glm(Shannon ~ phase, data = df, family = gaussian)))$p.value >= 0.05) {
  print("The model's errors are not significanty different from a normal distribution, continue with gaussian distribution")
} else {
  print("The model's errors are significanty different from a normal distribution, decide on another distribution")
}

(nema_sha_glm <- data.frame(tidy(glm(Shannon ~ phase, data = df, family = gaussian))) |> 
  mutate(group = "Nematodes", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3))) 

# we got a significant result, therefore we perform a pairwwise comparison test with emmeans
(nema_sha_pairwise <- data.frame(emmeans(glm(Shannon ~ phase, data = df, family = gaussian), 
                                specs = pairwise ~ phase)$contrast) |> 
  mutate(group = "Nematodes", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3)))


df |>
  mutate(
    phase = sub("Pioner", "P", phase),
    phase = sub("Bygg", "B", phase),
    phase = sub("Moden", "M", phase)
  ) |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Observed,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(color = "white"),
    axis.ticks.x = element_blank()
  ) +
  scale_fill_manual(values = c("#F7FBFF", "#9ECAE1", "#08519C")) +
  scale_color_manual(values = rep("black", times = length(unique(groundrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle("Nematodes") -> nemrich
nemrich

df |>
  mutate(
    phase = sub("Pioner", "P", phase),
    phase = sub("Bygg", "B", phase),
    phase = sub("Moden", "M", phase)
  ) |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Shannon,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_fill_manual(values = c("#F7FBFF", "#9ECAE1", "#08519C")) +
  scale_color_manual(values = rep("black", times = length(unique(groundrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle(" ") +
  geom_segment(
    x = 2, xend = 3,
    y = 2.3, yend = 2.3
  ) +
  annotate(
    geom = "text", label = "**", size = 6,
    x = 2.5, y = 2.35
  ) +
  coord_cartesian(ylim = c(0.8, 2.4)) -> nemsha
nemsha

ggarrange(nemrich, nemsha, nrow = 2) -> nem_alphaplot
nem_alphaplot
```

fungi
```{r}
# load and normalise phyloseq object:
fungi <- readRDS(here("data/clean_data/heathland-fire-fungi.RDS"))
fungi <- prune_taxa(taxa_sums(fungi) > 0, fungi)

funrar <- rarefy_even_depth(fungi,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = min(sample_sums(fungi))
)

# calculate alpha diversity measures and find medians and IQR:
(tibble(
  estimate_richness(funrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(funrar@sam_data)
) |>
  select(Observed, Shannon, phase) |>
  group_by(phase) |>
  mutate(
    group = "Fungi",
    N = vctrs::vec_count(phase)$count,
    q1_richness = quantile(Observed)[2],
    med_richness = quantile(Observed)[3],
    q3_richness = quantile(Observed)[4],
    q1_shannon = quantile(Shannon)[2],
    med_shannon = quantile(Shannon)[3],
    q3_shannon = quantile(Shannon)[4],
    richness = paste0(med_richness, " (IQR ", round(q3_richness-q1_richness, digits = 2), ")"), 
    shannon = paste0(round(med_shannon, digits = 2), " (IQR ", round(q3_shannon-q1_shannon, digits = 2), ")")
  ) |>
  select(group, phase, richness, shannon, N) |>
  distinct() |>
  arrange(phase = factor(
    phase,
    levels = c("P", "B", "M")
  )) -> alpha_stats_fungi)

# test whether differences in alpha diversity are significant between phases:
tibble(
  estimate_richness(funrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(funrar@sam_data)
) -> df

# richness

# we used GLM to test if richness was significantly different between successional phases. because richness is count data, a poisson distribution is appropriate. however, if the variance is larger than the mean richness, the data are overdispersed and we should use quasipoisson
if (var(df$Observed) > mean(df$Observed)) {
  print("Data is overdispersed, use quasipoisson distribution")
} else {
  print("Data is not overdispersed, use poisson distribution")
}

# run the GLM
(fun_rich_glm <- data.frame(tidy(glm(Observed ~ phase, data = df, family = quasipoisson))) |>
  mutate(p.value = round(p.value, digits = 3)) |> 
  mutate(group = "Fungi", .before = 1))

# we did not get a significant result, therefore we did not perform a pairwwise comparison 

# shannon diversity

# we test differences in Shannon diversity with a GLM. Shannon diversity is a continuous variable, so we basically want to use run the GLM with a gaussian distribution. we first check that the model's errors are not significantly different from a normal distribution using Shapiro-Wilk test

if (shapiro.test(rstandard(glm(Shannon ~ phase, data = df, family = gaussian)))$p.value >= 0.05) {
  print("The model's errors are not significanty different from a normal distribution, continue with gaussian distribution")
} else {
  print("The model's errors are significanty different from a normal distribution, decide on another distribution")
}

(fun_sha_glm <- data.frame(tidy(glm(Shannon ~ phase, data = df, family = gaussian))) |>
  mutate(p.value = round(p.value, digits = 3)) |> 
  mutate(group = "Fungi", .before = 1))

# we got a significant result, therefore we perform a pairwwise comparison test with emmeans
(fun_sha_pairwise <- data.frame(emmeans(glm(Shannon ~ phase, data = df, family = gaussian), 
                                specs = pairwise ~ phase)$contrast) |> 
  mutate(group = "Nematodes", .before = 1) |> 
  mutate(p.value = round(p.value, digits = 3)))


(df |>
  mutate(
    phase = sub("Pioner", "P", phase),
    phase = sub("Bygg", "B", phase),
    phase = sub("Moden", "M", phase)
  ) |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Observed,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(color = "white"),
    axis.ticks.x = element_blank()
  ) +
  scale_fill_manual(values = c("#DADAEB", "#9E9AC8", "#54278F")) +
  scale_color_manual(values = rep("black", times = length(unique(funrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle("Fungi") -> funrich)

(df |>
  mutate(
    phase = sub("Pioner", "P", phase),
    phase = sub("Bygg", "B", phase),
    phase = sub("Moden", "M", phase)
  ) |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Shannon,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_fill_manual(values = c("#DADAEB", "#9E9AC8", "#54278F")) +
  scale_color_manual(values = rep("black", times = length(unique(funrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle(" ") +
  geom_segment(
    x = 1, xend = 3,
    y = 4.85, yend = 4.85
  ) +
  annotate(
    geom = "text", label = ".", size = 6,
    x = 2, y = 4.95
  ) +
  geom_segment(
    x = 2, xend = 3,
    y = 4.7, yend = 4.7
  ) +
  annotate(
    geom = "text", label = ".", size = 6,
    x = 2.5, y = 4.8
  ) +
  coord_cartesian(ylim = c(3.8, 5)) -> funsha)

```

bacteria
```{r}
# load and normalise phyloseq object:
bacteria <- readRDS(here("data/clean_data/heathland-fire-bacteria.RDS"))

bacrar <- rarefy_even_depth(bacteria,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = 18963
)

# calculate alpha diversity measures and find medians and IQR:
(tibble(
  estimate_richness(bacrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(bacrar@sam_data)
) |>
  select(Observed, Shannon, phase) |>
  group_by(phase) |>
  mutate(
    group = "Bacteria",
    N = vctrs::vec_count(phase)$count,
    q1_richness = quantile(Observed)[2],
    med_richness = quantile(Observed)[3],
    q3_richness = quantile(Observed)[4],
    q1_shannon = quantile(Shannon)[2],
    med_shannon = quantile(Shannon)[3],
    q3_shannon = quantile(Shannon)[4],
    richness = paste0(med_richness, " (IQR ", round(q3_richness-q1_richness, digits = 2), ")"), 
    shannon = paste0(round(med_shannon, digits = 2), " (IQR ", round(q3_shannon-q1_shannon, digits = 2), ")")
  ) |>
  select(group, phase, richness, shannon, N) |>
  distinct() |>
  arrange(phase = factor(
    phase,
    levels = c("P", "B", "M")
  )) -> alpha_stats_bacteria)

# test whether differences in alpha diversity are significant between phases:
tibble(
  estimate_richness(bacrar,
    measures = c("Observed", "Shannon")
  ),
  data.frame(bacrar@sam_data)
) -> df

# richness

# we used GLM to test if richness was significantly different between successional phases. because richness is count data, a poisson distribution is appropriate. however, if the variance is larger than the mean richness, the data are overdispersed and we should use quasipoisson
if (var(df$Observed) > mean(df$Observed)) {
  print("Data is overdispersed, use quasipoisson distribution")
} else {
  print("Data is not overdispersed, use poisson distribution")
}

# run the GLM
(bac_rich_glm <- data.frame(tidy(glm(Observed ~ phase, data = df, family = quasipoisson))) |>
  mutate(p.value = round(p.value, digits = 3)) |> 
  mutate(group = "Bacteria", .before = 1))

# we did not get a significant result, therefore we did not perform a pairwwise comparison test 

# shannon diversity

# we test differences in Shannon diversity with a GLM. Shannon diversity is a continuous variable, so we basically want to use run the GLM with a gaussian distribution. we first check that the model's errors are not significantly different from a normal distribution using Shapiro-Wilk test

if (shapiro.test(rstandard(glm(Shannon ~ phase, data = df, family = gaussian)))$p.value >= 0.05) {
  print("The model's errors are not significanty different from a normal distribution, continue with gaussian distribution")
} else {
  print("The model's errors are significanty different from a normal distribution, decide on another distribution")
}

(bac_sha_glm <- data.frame(tidy(glm(Shannon ~ phase, data = df, family = gaussian))) |>
  mutate(p.value = round(p.value, digits = 3)) |> 
  mutate(group = "Bacteria", .before = 1))

# we did not get a significant result, therefore we did not perform a pairwwise comparison 

(df |>
  mutate(
    phase = sub("Pioner", "P", phase),
    phase = sub("Bygg", "B", phase),
    phase = sub("Moden", "M", phase)
  ) |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Observed,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(color = "white"),
    axis.ticks.x = element_blank()
  ) +
  scale_fill_manual(values = c("#FEE0D2", "#FC9272", "#CB181D")) +
  scale_color_manual(values = rep("black", times = length(unique(funrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle("Bacteria") -> bacrich)

(df |>
  mutate(
    phase = sub("Pioner", "P", phase),
    phase = sub("Bygg", "B", phase),
    phase = sub("Moden", "M", phase)
  ) |>
  ggplot(aes(
    x = factor(phase, levels = c("P", "B", "M")),
    y = Shannon,
    fill = factor(phase, levels = c("P", "B", "M")),
    color = factor(phase, levels = c("P", "B", "M"))
  )) +
  theme_classic() +
  geom_boxplot(linewidth = 0.6, alpha = 1, outlier.shape = NA, width = 0.5) +
  geom_jitter(
    size = 2, shape = 21, stroke = 1, alpha = 0.5,
    position = position_jitter(height = 0, width = 0.3)
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_fill_manual(values = c("#FEE0D2", "#FC9272", "#CB181D")) +
  scale_color_manual(values = rep("black", times = length(unique(funrar@sam_data$phase)))) +
  ylab(" ") +
  ggtitle("Bacteria") -> bacsha)

```

combine plots into figure 2(?)
```{r}
ggarrange(plants_alphaplot, malaise_alphaplot, ground_alphaplot, nem_alphaplot, fun_alphaplot, bac_alphaplot,
  nrow = 1
)
```


make supplementary table 1
```{r}
(rbind(
  alpha_stats_plants, 
  alpha_stats_malaise, 
  alpha_stats_pitfall, 
  alpha_stats_nematodes, 
  alpha_stats_fungi, 
  alpha_stats_bacteria
))
```

