---
title: "taxonomy"
author: "pgfarsund"
date: "`r Sys.Date()`"
output: html_document
---

# we will summarise the relative abundance of (relatively) high taxonomic levels 
# for each group and across the successional stages. the three top taxa should
# be mentioned in the text and will be accompanied by krona charts. metacoder
# trees will also be included, but they should emphasise differentially 
# abundant taxa rather than taxonomic composition. 

# load library
```{r}
library(here)
library(phyloseq)
library(tidyverse)
library(psadd)
library(ggpubr)
#
```


function returning plot
```{r}
plot_top_relab <- function(ps, # a phyloseq object
                           taxlevel, # the taxonomic level you want to plot
                           how_many_taxa, # how many taxa should not be grouped as "Other"?
                           group # set to "phase"
) {
  # load a few packages
  library(tidyverse)
  library(RColorBrewer)

  # use this object to calculate relative abundance at the phylum level
  glom <- tax_glom(ps, taxrank = taxlevel, NArm = F)

  # record the names of the top ntaxa taxa
  top <- names(sort(taxa_sums(glom), decreasing = TRUE))[1:how_many_taxa]

  # remove taxa that are not among the top ntaxa taxa
  top <- prune_taxa(top, glom)

  # remove the top ntaxa
  bottom <- names(sort(taxa_sums(glom), decreasing = FALSE))[1:c(ntaxa(glom) - ntaxa(top))]

  # remove the top ntaxa taxa
  bottom <- prune_taxa(bottom, glom)

  # group bottom taxa as "Other"
  bottom@tax_table[, 1:ncol(bottom@tax_table)] <- "Other"

  # agglomorate bottom taxa
  bottom <- tax_glom(bottom, taxrank = taxlevel, NArm = F)

  # merge the top and bottom objects
  top <- merge_phyloseq(top, bottom)

  # melt the ps object to a data frame
  df <- psmelt(top) %>%
    rename("taxlevel" = taxlevel)

  # replace NA's with "Unassigned"
  # df$taxlevel[is.na(df$taxlevel)] <- "Unassigned"

  # make a custom color palette with ntaxa colors + grey for "Other":
  custcol <- rep(brewer.pal(12, "Paired"), times = 100)
  # custcol <- c(custcol[1:how_many_taxa+1], "grey")

  # make stuff for fill:
  fillvector <- data.frame(df$taxlevel, df$Abundance) |>
    filter(df.taxlevel != "Other")

  fillvector <- aggregate(fillvector$df.Abundance,
    by = list(fillvector$df.taxlevel), FUN = sum
  )

  colnames(fillvector) <- c("df.taxlevel", "df.Abundance")

  # we use the following vector to sort the taxa in barplot
  fillvector <- fillvector[order(fillvector$df.Abundance, decreasing = TRUE), ]
  fillvector <- unique(fillvector$df.taxlevel)

  # now we calculate 100 % relative abundance for each group individually
  df |>
    reframe(abundance = sum(Abundance), .by = c(phase, taxlevel)) |>
    group_by(phase) |>
    mutate(relab = abundance / sum(abundance)) |>
    ungroup() |>
    mutate(test = sum(relab)) |>
    mutate(phase = case_match(
      phase,
      c("P", "Pioner", "Pioneer") ~ "P",
      c("B", "Bygg") ~ "B",
      c("M", "Moden") ~ "M"
    )) |>
    # now we can make the figure
    ggplot(aes(
      x = factor(phase, levels = c("M", "B", "P")),
      y = relab,
      fill = factor(taxlevel, levels = c(fillvector, "Other")),
      color = factor(taxlevel, levels = c(fillvector, "Other"))
    )) +
    # theme_grey(base_size = 18) +
    theme_grey() +
    geom_col(width = 0.75) +
    scale_fill_manual(values = custcol) +
    scale_color_manual(values = rep("black", times = how_many_taxa + 1)) +
    scale_y_continuous(labels = scales::percent_format()) +
    guides(
      fill = guide_legend(title = taxlevel, nrow = 4),
      color = guide_legend(title = taxlevel, nrow = 4)
    ) +
    theme(
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.text = element_text()
    ) +
    coord_flip()
}
```


function returning dataframe
```{r}
get_top_relab <- function(ps, # a phyloseq object
                          taxlevel, # the taxonomic level you want to plot
                          how_many_taxa, # how many taxa should not be grouped as "Other"?
                          group # should the data be grouped by a variable?
) {
  # load a few packages
  library(tidyverse)
  library(RColorBrewer)

  # use this object to calculate relative abundance at the phylum level
  glom <- tax_glom(ps, taxrank = taxlevel, NArm = F)

  # record the names of the top ntaxa taxa
  top <- names(sort(taxa_sums(glom), decreasing = TRUE))[1:how_many_taxa]

  # remove taxa that are not among the top ntaxa taxa
  top <- prune_taxa(top, glom)

  # remove the top ntaxa
  bottom <- names(sort(taxa_sums(glom), decreasing = FALSE))[1:c(ntaxa(glom) - ntaxa(top))]

  # remove the top ntaxa taxa
  bottom <- prune_taxa(bottom, glom)

  # group bottom taxa as "Other"
  bottom@tax_table[, 1:ncol(bottom@tax_table)] <- "Other"

  # agglomorate bottom taxa
  bottom <- tax_glom(bottom, taxrank = taxlevel, NArm = F)

  # merge the top and bottom objects
  top <- merge_phyloseq(top, bottom)

  # melt the ps object to a data frame
  df <- psmelt(top) %>%
    rename("taxlevel" = taxlevel) |>
    reframe(abundance = sum(Abundance), .by = c(phase, taxlevel)) |>
    group_by(phase) |>
    mutate(
      relative_abundance = abundance / sum(abundance),
      relative_abundance = round(relative_abundance, digits = 4)
    ) |>
    ungroup() |>
    mutate(phase = case_match(
      phase,
      c("P", "Pioner", "Pioneer") ~ "P",
      c("B", "Bygg") ~ "B",
      c("M", "Moden") ~ "M"
    )) |>
    arrange(
      phase, -relative_abundance
    )
}
```


bacteria
```{r}
bacteria <- readRDS(here("data/clean_data/heathland-fire-bacteria.RDS"))
bacteria

# normalize the number of reads per sample:
bacrar <- rarefy_even_depth(bacteria,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = 18963
)
bacteria
bacrar

bactaxplot <- plot_top_relab(
  ps = bacrar, taxlevel = "Phylum",
  how_many_taxa = 10, group = "phase"
) +
  ggtitle("Bacteria")

data.frame(get_top_relab(
  ps = bacrar,
  taxlevel = "Phylum",
  how_many_taxa = 10,
  group = "phase"
)) |>
  select(-abundance) |> 
  arrange(factor(phase, levels = c("P", "B", "M")), 
          -relative_abundance)
```


fungi
```{r}
fungi <- readRDS(here("data/clean_data/heathland-fire-fungi.RDS"))
fungi <- prune_taxa(taxa_sums(fungi) > 0, fungi)
sample_names(fungi) <- paste0("sample_", 1:nsamples(fungi))

# fix NA's
tax <- data.frame(fungi@tax_table) %>%
  replace(is.na(.), values = "Unassigned") %>%
  mutate(
    Kingdom = sub("k__", "", Kingdom),
    Phylum = sub("p__", "", Phylum),
    Class = sub("c__", "", Class)
  )

# put back in ps object
fungi@tax_table <- tax_table(as.matrix(tax))

# normalize the number of reads per sample
funrar <- rarefy_even_depth(fungi,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = min(sample_sums(fungi))
)

# make relative abundance figure
funtaxplot <- plot_top_relab(
  ps = funrar, taxlevel = "Class",
  how_many_taxa = 11, group = "phase"
) +
  ggtitle("Fungi")
funtaxplot

data.frame(get_top_relab(
  ps = funrar,
  taxlevel = "Class",
  how_many_taxa = 11,
  group = "phase"
)) |>
  select(-abundance) |> 
  arrange(factor(phase, levels = c("P", "B", "M")), 
          -relative_abundance)
```


nematodes
```{r}
nematodes <- readRDS(here("data/clean_data/heathland-fire-nematodes.RDS"))
nematodes <- prune_taxa(taxa_sums(nematodes) > 0, nematodes)

nemrar <- rarefy_even_depth(nematodes,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = min(sample_sums(nematodes))
)
nematodes
nemrar

nemtaxplot <- plot_top_relab(
  ps = nemrar,
  taxlevel = "genus",
  how_many_taxa = 10,
  group = "phase"
) +
  ggtitle("Nematodes") +
  guides(
    fill = guide_legend(title = "Genus", ncol = 3, nrow = 4),
    color = guide_legend(title = "Genus", ncol = 3, nrow = 4)
  )

data.frame(get_top_relab(
  ps = nemrar,
  taxlevel = "genus",
  how_many_taxa = 10,
  group = "phase"
)) |>
  select(-abundance) |> 
  arrange(factor(phase, levels = c("P", "B", "M")), 
          -relative_abundance)
```


walking arthropods
```{r}
pit <- readRDS(here("data/clean_data/heathland-fire-pitfall.RDS"))
pit <- prune_taxa(taxa_sums(pit) > 0, pit)

pitrar <- rarefy_even_depth(pit,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = 10
)

pittaxplot <- plot_top_relab(
  ps = pitrar, taxlevel = "Order",
  how_many_taxa = 10, group = "phase"
) +
  ggtitle("Pitfall traps") +
  guides(
    fill = guide_legend(title = "Order", ncol = 3, nrow = 4),
    color = guide_legend(title = "Order", ncol = 3, nrow = 4)
  )

data.frame(get_top_relab(
  ps = pitrar,
  taxlevel = "Order",
  how_many_taxa = 10,
  group = "phase"
)) |>
  select(-abundance) |> 
  arrange(factor(phase, levels = c("P", "B", "M")), 
          -relative_abundance)
```


flying insects
```{r}
insects <- readRDS(here("data/insects/heathland-fire-insects.RDS"))
insects
insects <- prune_taxa(taxa_sums(insects) > 0, insects)
insects

inrar <- rarefy_even_depth(insects,
  rngseed = 123,
  replace = F,
  trimOTUs = T,
  sample.size = min(sample_sums(insects))
)

(intaxplot <- plot_top_relab(
  ps = inrar, taxlevel = "order",
  how_many_taxa = 10, group = "phase"
) +
  ggtitle("Malaise traps") +
  coord_flip(ylim = c(0, 0.425)) +
  guides(
    fill = guide_legend(title = "Order", ncol = 3, nrow = 4),
    color = guide_legend(title = "Order", ncol = 3, nrow = 4)
  ))

relab <- get_top_relab(
  ps = inrar, taxlevel = "order",
  how_many_taxa = 10,
  group = "phase"
)
relab
```


plants
```{r}
plants <- readRDS("data/plants/heathland-fire-plants.RDS")
plants <- prune_taxa(taxa_sums(plants) > 0, plants)
plants

# make plot
(plataxplot <- plot_top_relab(
  ps = plants, taxlevel = "species",
  how_many_taxa = 10, group = "phase"
) +
  ggtitle("Plants") +
  guides(
    fill = guide_legend(title = "Species", ncol = 3, nrow = 4, ),
    color = guide_legend(title = "Species", ncol = 3, nrow = 4)
  ))

# get relative abundances
relab <- get_top_relab(
  ps = plants, taxlevel = "functional_groups",
  how_many_taxa = ntaxa(plants),
  group = "phase"
)
relab
```


arrange the plots
```{r}
library(ggpubr)

ggarrange(
  plataxplot,
  intaxplot,
  wataxplot,
  nemtaxplot,
  funtaxplot,
  bactaxplot,
  nrow = 2,
  ncol = 3,
  widths = c(
    1, 1, 1,
    1, 1, 1
  ),
  heights = c(
    1, 1, 1,
    1, 1, 1
  )
)
# 7 x 17.5
```
